IRS County-to-County Migration Data

### 2018-08-16 manuscript_file.Rmd
organization.

---
output: 
  pdf_document:
    number_sections: true
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: LATEX/svm-latex-ms.tex
title: "IRS County-to-County Migration Data"
thanks: "The data and code that supports this analysis are available in the supplementary materials."
author:
  - name: Mathew E. Hauer ^1,2^*
    affiliation: Florida State University
# abstract: ""
keywords: "Migration, IRS, Data, R"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
#fontfamily: mathpazo
fontsize: 12pt
spacing: single
bibliography:  LATEX/mybibfile.bib
biblio-style: apsr
# use apsr or nature
header-includes:
- \usepackage[all]{nowidow}
- \usepackage{rotating}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead{}
- \fancyhead[CE,CO]{IRS Migration Data}

---
<!-- *Corresponding author. mehauer\@fsu.edu. p: 850-644-7103 -->

@@

@@

**CONTRIBUTION**: To encourage and facilitate the use of this data, I provide the full R script to download and flatten the IRS migration data as counts and a finalized data set covering the period 1990-2010.

@@

To encourage and facilitate the use of this tremendous migration resource, I flattened the IRS migration data into a single, standardized file containing all migration flows for the period 1990-2010. By publishing both the R code to collate, wrangle, and flatten the IRS migration data as well as the migration data itself, my hope is to save time for other migration scholars and facilitate their use. Scholars who wish to use these data should still familiarize themselves with the strengths and weaknesses, idosyncracies, and design of these data (see [@gross2005internal; @engels1981measuring;@franklin2006pandora;@pierce2015soi] and with the procedures outlined in this document and in the corresponding R code.

While the IRS migration data presently continues to 2015, I limit the data to the period 1990-2010. Starting in tax year 2012 (migration year 2011), the IRS introduced a change in their processing [@pierce2015soi] creating a break in the time series.

@@

The differences described above and in **Table 1** are only some of the differences that are of interest to the data I produce here. Total Migrants, ie FIPS 96000 for migration data after 1995, is also broken down into Total Mig - US (FIPS 97000), Total Mig - US Same State (FIPS 97001), Total Mig - US Diff St (FIPS 97003), and Total Mig - Foreign (FIPS 98000). The IRS did not code these migration flows in this manner for all years, and in some cases (such as Total Mig - Foreign) migration flows are not reported. For simplicity and data continuity purposes, I simply create a new origin/destination (FIPS 99999) that contains all unspecified migration flows. I do this by subtracting the number of migrants from migration flows with greater than 10 migrants from the total number of migrants. This way, the sum of all migrants will equal the total number of migrants.

Besides the aggregation to FIPS 99999, I performed no other post-processing of the IRS data.

@@

## Setup 
The first part of the code simply sets up the R workspace to facilitate the data wrangling. The approriate R packages are downloaded and installed if the user does not already have these packages installed. The paralell computing environment is also set up as `DetectCores() - 1` to ensure one user core is left for other tasks that computer might need. A single reference tab seperated (tsv) file is required in this section. `ref_state.tsv` contains FIPS code information for US states. I simply add an additional FIPS code for 'unknown' and assign it FIPS state 99.

@@

US Counties are fairly stable geographic units but some changes in counties do exist. To try and keep as close to the original data fidelity as possible, I did not recode any geographic changes. For instance, Broomfield County CO (FIPS 07014) was created out of parts of Adams, Boulder, Jefferson, and Weld counties and thus contains data only after 2002. Users should be aware of any changes in geography that could substantially alter any analyses^[More detailed information about county boundary changes can be found at the following locations https://www.census.gov/geo/reference/county-changes.html
http://www.nber.org/asg/ASG_release/County_City/FIPS/FIPS_Changes.pdf
https://www.cdc.gov/nchs/data/nvss/bridged_race/County_Geography_Changes.pdf
https://www.ddorn.net/data/FIPS_County_Code_Changes.pdf].

Users should be aware that I prepared these data to stay as close to the original data fidelity as possible, documented their creation, and with open-source computer code. These data should be used onlly with full awareness of the inherent limitations of the IRS migration data and with the knowledge of the procedures outlined in this document and in the corresponding R code.

| origin    | destination | 1990          | 1991          | 1992                           | ...                                | 2010  |
|-----------|-------------|---------------|---------------|--------------------------------|------------------------------------|-------|
| 01001     | 01001       | 10785         | 11010         | 11196                          | ...                                | 17278 |
| 01001     | 01003       | 0             | 0             | 12                             | ...                                | 11    |
| ...       | ...         | ...           | ...           | ...                            | ...                                | ...   |
| 01001     | 99999       | 524           | 403           | 479                            | ...                                | 732   |
Table: Selected file format for the final flat file. Origin and Destinations are the five-digit FIPS codes with 99999 representing all flows with fewer than 10 filers.

>>>>

---
output: 
  pdf_document:
    number_sections: true
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: LATEX/svm-latex-ms.tex
title: "IRS County-to-County Migration Data"
thanks: "The data and code that supports this analysis are available in the supplementary materials."
author:
  - name: Mathew E. Hauer ^1,2^*
    affiliation: Florida State University
  - name: James M. Byars ^3^*
    affiliation: University of Georgia
# abstract: ""
keywords: "Migration, IRS, Data, R"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
#fontfamily: mathpazo
fontsize: 12pt
spacing: single
bibliography:  LATEX/mybibfile.bib
biblio-style: apsr
# use apsr or nature
header-includes:
- \usepackage[all]{nowidow}
- \usepackage{rotating}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead{}
- \fancyhead[CE,CO]{IRS Migration Data}

---
*Corresponding author. mehauer\@fsu.edu. p: 850-644-7103

@@

^3^ Carl Vinson Institute of Government, University of Georgia.

@@

**CONTRIBUTION**: To encourage and facilitate the use of this data, we provide the full R script to download and flatten the IRS migration data as counts and a finalized data set covering the period 1990-2010.

@@

To encourage and facilitate the use of this tremendous migration resource, we flattened the IRS migration data into a single, standardized file containing all migration flows for the period 1990-2010. By publishing both the R code to collate, wrangle, and flatten the IRS migration data as well as the migration data itself, my hope is to save time for other migration scholars and facilitate their use. Scholars who wish to use these data should still familiarize themselves with the strengths and weaknesses, idosyncracies, and design of these data (see [@gross2005internal; @engels1981measuring;@franklin2006pandora;@pierce2015soi] and with the procedures outlined in this document and in the corresponding R code.

While the IRS migration data presently continues to 2015, we limit the data to the period 1990-2010. Starting in tax year 2012 (migration year 2011), the IRS introduced a change in their processing [@pierce2015soi] creating a break in the time series.

@@

The differences described above and in **Table 1** are only some of the differences that are of interest to the data we produce here. Total Migrants, ie FIPS 96000 for migration data after 1995, is also broken down into Total Mig - US (FIPS 97000), Total Mig - US Same State (FIPS 97001), Total Mig - US Diff St (FIPS 97003), and Total Mig - Foreign (FIPS 98000). The IRS did not code these migration flows in this manner for all years, and in some cases (such as Total Mig - Foreign) migration flows are not reported. For simplicity and data continuity purposes, we simply create a new origin/destination (FIPS 99999) that contains all unspecified migration flows. we do this by subtracting the number of migrants from migration flows with greater than 10 migrants from the total number of migrants. This way, the sum of all enumerated migrants in my dataset will equal the total number of migrants in the IRS dataset.

The aggregation to FIPS 99999 is the only post-processing of the IRS data.

@@

## Setup 
The first part of the code simply sets up the R workspace to facilitate the data wrangling. The approriate R packages are downloaded and installed if the user does not already have these packages installed. The paralell computing environment is also set up as `DetectCores() - 1` to ensure one user core is left for other tasks that computer might need. A single reference tab seperated (tsv) file is required in this section. `ref_state.tsv` contains FIPS code information for US states. we simply add an additional FIPS code for 'unknown' and assign it FIPS state 99.

@@

US Counties are fairly stable geographic units but some changes in counties do exist. To try and keep as close to the original data fidelity as possible, we did not recode any geographic changes. For instance, Broomfield County CO (FIPS 07014) was created out of parts of Adams, Boulder, Jefferson, and Weld counties and thus contains data only after 2002. Users should be aware of any changes in geography that could substantially alter any analyses^[More detailed information about county boundary changes can be found at the following locations https://www.census.gov/geo/reference/county-changes.html
http://www.nber.org/asg/ASG_release/County_City/FIPS/FIPS_Changes.pdf
https://www.cdc.gov/nchs/data/nvss/bridged_race/County_Geography_Changes.pdf
https://www.ddorn.net/data/FIPS_County_Code_Changes.pdf].

Users should be aware that we prepared these data to stay as close to the original data fidelity as possible, documented their creation, and with open-source computer code. These data should be used onlly with full awareness of the inherent limitations of the IRS migration data and with the knowledge of the procedures outlined in this document and in the corresponding R code.

| origin    | destination | 1990          | 1991          | 1992                           | ...                                | 2010  |
|-----------|-------------|---------------|---------------|--------------------------------|------------------------------------|-------|
| 01001     | 01001       | 26703         | 27278         | 28677                          | ...                                | 40643 |
| 01001     | 01003       | 0             | 0             | 27                             | ...                                | 39    |
| ...       | ...         | ...           | ...           | ...                            | ...                                | ...   |
| 01001     | 99999       | 1324           | 1020           | 1200                            | ...                                | 1758   |
Table: Selected file format for the final flat file. Origin and Destinations are the five-digit FIPS codes with 99999 representing all flows with fewer than 10 filers.
